# Import Data {#import_data}

```{r, include=FALSE, warning=F, message=F}
# data mgmt
library(tidyverse)
library(lubridate)
# visualization
library(RColorBrewer)
library(scales)
library(viridis)
library(kableExtra)
# spatial
library(sf)
library(stars)
library(starsExtra) # functions for working with stars objects
library(lwgeom) 
library(mapview) #Interactive maps
library(leafpop) #map html popup
library(elevatr) # elevation data (DEMs)
library(pracma) # Practical Numerical Math Functions

# knit options
knitr::opts_chunk$set(
  echo = TRUE
  , warning = FALSE
  , message = FALSE
  , results='hide'
  , fig.width = 10
  , fig.height = 7
)
# qualitative color pallette
my_color_pal_qual <- function(n) viridis::turbo(n = n * 4)[which(seq(1,n * 4,1) %% 4 == 0)]
```

```{r}
# turn off the s2 processing 
## https://stackoverflow.com/questions/68478179/how-to-resolve-spherical-geometry-failures-when-joining-spatial-data
sf::sf_use_s2(FALSE)

# set crs
my_crs <- 5070 # EPSG:5070 = NAD83/Conus Albers (units = meters); EPSG:4326 = WGS 84 (units = dd); EPSG:4269 = NAD83 (units = dd)
```

## National Forest Management data download

The Forest Activity Tracking System (FACTS) [database](https://data.fs.usda.gov/geodata/edw/datasets.php?xmlKeyword) maintained by the U.S. Department of Agriculture, Forest Service (USFS) includes georeferenced boundaries of national forests.

```{r}
# load boundary shapefile
  forests <- sf::st_read("../data/forests.gpkg") %>%
    rename_with(~ tolower(
      gsub(" ", "_",
         str_trim(gsub("\\s+", " ", .x))
      )
    )) %>% 
    sf::st_transform(my_crs) # EPSG:4326 = WGS 84
# keep GMUG
gmug <- forests %>% 
  dplyr::filter(
    forest_commonname %in% c(
      "Grand Mesa National Forest"
      , "Uncompahgre National Forest"
      , "Gunnison National Forest"
    )
  ) %>% 
  sf::st_union()

```

### GMUG Map

```{r, results='asis'}
mapview::mapviewOptions(homebutton = FALSE, basemaps = c("Esri.WorldTopoMap")) # "Esri.WorldImagery"
mapview::mapview(
  x = gmug
  , color = "black"
  , lwd = 4
  , alpha.regions = 0
  , label = FALSE
  , legend = FALSE
  , popup = FALSE
  , map.types = "Esri.WorldTopoMap" # "Esri.WorldImagery"
)
```


## Load Treatment data

```{r}
############################
# load shapefiles
############################
shp_import_fn <- function(pth){
  # import
  dta <- sf::st_read(pth) %>% 
    dplyr::filter(sf::st_is_valid(.)) %>% 
    rename_with(~ tolower(
      gsub(" ", "_", 
         str_trim(gsub("\\s+", " ", .x))
      )
    )) %>% 
    sf::st_transform(crs = sf::st_crs(forests)) %>% 
    dplyr::mutate(area_m2 = as.numeric(sf::st_area(.)))
  #rename sf geom column
  names(dta)[names(dta)==tolower(attr(dta, "sf_column"))] = "geometry"
  sf::st_geometry(dta) = "geometry"
  # return
  return(dta)
}

# group selection openings
openings <- shp_import_fn("../data/Bald_all_groups.shp")

# group selection reserve groups
reserves <- shp_import_fn("../data/Bald_all_Reserve_grps.shp")

# harvest units
units <- shp_import_fn("../data/UnitEstimate.shp")
```

### GMUG and treatment area map

```{r}
ggplot() + 
  geom_sf(
    data = gmug
    , alpha = 0
    , lwd = 1, color = "black" 
  ) +
  geom_sf(
    data = units %>% sf::st_union() %>% sf::st_buffer(dist = 500)
    , fill = "royalblue"
    , lwd = NA
  ) +
  labs(
    title = "GMUG boundary and treatment area in blue"
  ) +
  theme_bw() + 
  theme(
    legend.position = "none"
    , axis.text.y = element_text(size = 7)
    , axis.text.x = element_text(size = 7, angle = 35, hjust = 0.7)
    , panel.border = element_blank()
  ) 
```

### Treatment units and group map

```{r}
ggplot() + 
  geom_sf(
    data = units
    , mapping = aes(fill = unit_numbe)
    , lwd = 0.8, color = "black" 
  ) +
  geom_sf(data = openings, fill = "white", color = "firebrick", lwd = 1) +
  geom_sf(data = reserves, fill = "black", color = "black") +
  scale_fill_viridis_d(option = "turbo", alpha = 0.4) +
  labs(
    title = "Harvest treatment units and groups"
    , subtitle = "Reserves = black, Openings = white"
  ) +
  theme_bw() + 
  theme(
    legend.position = "none"
    , axis.text.y = element_text(size = 7)
    , axis.text.x = element_text(size = 7, angle = 35, hjust = 0.7)
    , panel.border = element_blank()
  ) 
```


## Load Elevation Data

The function `elevatr::get_elev_raster` was used to obtain a digital elevation model (DEM) raster (~6.8m resolution).

```{r}
#######################################################
#######################################################
# read elevation data
#######################################################
#######################################################
  if(file.exists("../data/elev.tif") == FALSE){
      # z =14 is highest resolution (~6.8m)
      elev <- elevatr::get_elev_raster(
          locations = units %>% 
            sf::st_union() %>% 
            sf::st_buffer(dist = 100)
          , z = 14
        ) %>% 
        stars::st_as_stars() %>% 
        setNames("elev")
      # save
      stars::write_stars(elev, "../data/elev.tif", append = FALSE)
  }else{
    elev <- stars::read_stars("../data/elev.tif")
  }
```

### Elevation plot

```{r, warning=F, message=F, fig.width = 10, fig.height = 6}
# plot
ggplot() + stars::geom_stars(data = elev[
    units %>% 
      sf::st_union() %>% 
      sf::st_buffer(dist = 200) %>% 
      sf::st_bbox()
  ]) +
  scale_fill_viridis_c(option = "viridis", alpha = 0.9, na.value = "transparent") +
  geom_sf(data = units, alpha = 0, lwd = 1, color = "black") +
  labs(
      title = "GMUG Treatment Units Elevation Map"
      , subtitle = sf::st_crs(elev)$input
    ) +
  xlab("") +
  ylab("") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme_bw() +
  theme(
    legend.position = "bottom"
    , legend.direction = "horizontal"
    , legend.margin = margin(0,0,0,0)
    , legend.box.margin = margin(-21,0,2,0)
    , legend.text = element_text(size = 7)
    , legend.title = element_text(size = 7)
    , axis.text.y = element_text(size = 7)
    , axis.text.x = element_text(size = 7, angle = 35, hjust = 0.7)
    , panel.border = element_blank()
  ) +
  guides(
    fill = guide_legend(title="Elev. (m)")
  )

```

```{r, echo=FALSE, include=FALSE}
gc()
```

## Aspect and HLI

Aspect of a slope (the direction or azimuth that a slope faces) strongly influences potential direct incident radiation and temperature. Untransformed, aspect is a poor variable for quantitative analysis, since 1° is adjacent to 360° – the numbers are very different even though the aspect is about the same. The following equation for Heat Load Index (HLI) rescales aspect to a scale of zero to one, with zero being the coolest slope (northeast) and one being the warmest slope (southwest). HLI is a proxy for aspect-driven solar heating.

$$
\textrm{Heat load index} = \frac{1 - \cos(\theta - 45)}{2}
$$
where $\theta$ = aspect in degrees east of north. While this and related equations are useful, they do not consider the steepness of the slope. For example, a 1° south-facing slope would receive the same heat load index as a 30° south-facing slope, even though the latter will be considerably warmer.

HLI calculations have been developed which combine slope, folded aspect (i.e., “southwestness”), and latitude to estimate terrain-driven solar heating at a given location ([McCune & Keon 2002: eq. 2](https://onlinelibrary.wiley.com/doi/10.1111/j.1654-1103.2002.tb02087.x)). To calculate HLI, we used equation 2 of [McCune & Keon 2002: eq. 2](https://onlinelibrary.wiley.com/doi/10.1111/j.1654-1103.2002.tb02087.x). Following calculation, we aggregated these data to each group selection boundary (median HLI). Higher values of HLI are typical on steep southwesterly slopes, while low values are common on steep northeasterly slopes. Moderate values are common in flatter terrain and on northwesterly and southeasterly aspects. 

Aspect must first be "folded" about the north-south line, rescaling 0-360° to 0-180°, such that NE = NW, E = W, etc: 

$$
\textrm{Folded aspect°} = 180° \, - \, | Aspect° – 180° |
$$
```{r}
# calculate slope and aspect from DEM
topo <- c(
    elev
    , starsExtra::slope(elev)
    , starsExtra::aspect(elev)
  ) %>% 
  setNames(
    c("elev", "slope_deg", "aspect_deg")
  ) %>% 
  dplyr::mutate(
    slope_rad = pracma::deg2rad(as.numeric(slope_deg))
    , aspect_rad = pracma::deg2rad(as.numeric(aspect_deg))
    , folded_aspect_deg = 180 - abs(as.numeric(aspect_deg)-180)
    , folded_aspect_rad = pracma::deg2rad(folded_aspect_deg)
  )
# map
```

### Topography maps

```{r}
vars <- c(
  "elev"
  , "slope_deg"
  , "aspect_deg"
  , "folded_aspect_deg"
)
vars_title <- c(
  "Elevation (m)"
  , "Slope (\u00B0)"
  , "Aspect (\u00B0)"
  , "Folded Aspect (\u00B0)"
)
vars_color <- c(
  "viridis"
  , "cividis"
  , "rocket"
  , "mako"
)
plts <- list()
for(x in 1:length(vars)){
  plts[[x]] <- 
      ggplot() + stars::geom_stars(data = (topo %>% dplyr::select(vars[x]))[
        units %>% 
          sf::st_union() %>% 
          sf::st_buffer(dist = 200) %>% 
          sf::st_bbox()
      ]) +
      scale_fill_viridis_c(option = vars_color[x], alpha = 0.9, na.value = "transparent") +
      geom_sf(data = units, alpha = 0, lwd = 1, color = "black") +
      labs(
          title = vars_title[x]
        ) +
      xlab("") +
      ylab("") +
      scale_x_continuous(expand = c(0, 0)) +
      scale_y_continuous(expand = c(0, 0)) +
      theme_bw() +
      theme(
        legend.position = "bottom"
        , legend.direction = "horizontal"
        , legend.margin = margin(0,0,0,0)
        , legend.box.margin = margin(-21,0,2,0)
        , legend.text = element_text(size = 7)
        , legend.title = element_text(size = 7)
        , axis.text.y = element_text(size = 7)
        , axis.text.x = element_text(size = 7, angle = 35, hjust = 0.7)
        , panel.border = element_blank()
      ) +
      guides(
        fill = guide_legend(title=vars_title[x])
      )
}

# combine plots
cowplot::plot_grid(plotlist =  plts, nrow = 2, ncol = 2)
```



```{r}
# functions to aggregate raster data to vector data
topo_calc_fn <- function(dta){
  # calculate median aspect and slope by vector
  temp_slope <- aggregate(topo %>% dplyr::select(slope_rad), dta, median) 
  temp_aspect <- aggregate(topo %>% dplyr::select(aspect_rad), dta, median) 
  temp_folded_aspect <- aggregate(topo %>% dplyr::select(folded_aspect_rad), dta, median)
  # vector center
  temp_center <- sf::st_centroid(dta)$geometry %>% 
    sf::st_transform(crs = 4326) %>% 
    sf::st_coordinates()
  # attach to input data
  new_dta <- dta %>% 
    dplyr::mutate(
      slope_rad = temp_slope$slope_rad
      , aspect_rad = temp_aspect$aspect_rad
      , folded_aspect_rad = temp_folded_aspect$folded_aspect_rad
      , latitude_rad = pracma::deg2rad(as.numeric(temp_center[,2]))
      , slope_deg = pracma::rad2deg(temp_slope$slope_rad)
      , aspect_deg = pracma::rad2deg(temp_aspect$aspect_rad)
      , folded_aspect_deg = pracma::rad2deg(temp_folded_aspect$folded_aspect_rad)
      , latitude_deg = as.numeric(temp_center[,2])
    )
  return(new_dta)
}
# apply to data
# openings <- openings %>% topo_calc_fn()
# reserves <- reserves %>% topo_calc_fn()



```

```{r}
# hey
```

