# Analysis of HLI {#analysis}

```{r, include=FALSE, warning=F, message=F}
# data mgmt
library(tidyverse)
library(lubridate)
# visualization
library(RColorBrewer)
library(scales)
library(viridis)
library(kableExtra)
library(latex2exp)
# spatial
library(sf)
library(stars)
library(starsExtra) # functions for working with stars objects
library(lwgeom) 
library(mapview) #Interactive maps
library(leafpop) #map html popup

# knit options
knitr::opts_chunk$set(
  echo = TRUE
  , warning = FALSE
  , message = FALSE
  , results='hide'
  , fig.width = 10
  , fig.height = 7
)
# qualitative color pallette
my_color_pal_qual <- function(n) viridis::turbo(n = n * 4)[which(seq(1,n * 4,1) %% 4 == 0)]
```

```{r}
# turn off the s2 processing 
## https://stackoverflow.com/questions/68478179/how-to-resolve-spherical-geometry-failures-when-joining-spatial-data
sf::sf_use_s2(FALSE)
```


## Load Data

Spatial data was loaded and cleaned in [prior chapter](#data_prep). 

```{r}
# read
topo <- readRDS("../data/topo.RDS")
all_groups_topo <- sf::st_read("../data/all_groups_topo.gpkg")
units <- sf::st_read("../data/units.gpkg")
```

## Set up base map

```{r}
# map
vars <- c(
  "elev"
  , "slope_deg"
  , "aspect_deg"
  , "folded_aspect_deg"
  , "hli"
)
vars_title <- c(
  "Elevation (m)"
  , "Slope (\u00B0)"
  , "Aspect (\u00B0)"
  , "Folded Aspect (\u00B0)"
  , "Heat Load Index (HLI)"
)
vars_color <- c(
  "viridis"
  , "cividis"
  , "inferno"
  , "plasma"
  , "turbo"
)
vars_llim <- c(
  min(topo$elev, na.rm=T)*.95
  , 0
  , 0
  , 0
  , 0
)
vars_ulim <- c(
  max(topo$elev, na.rm=T)*1.05
  , max(as.numeric(topo$slope_deg), na.rm=T)*1.05
  , 360
  , 180
  , 1
)
# plts <- list()
plt_fn <- function(x){
  (
      ggplot() + 
      stars::geom_stars(data = (topo %>% dplyr::select(vars[x]))[]) +
      scale_fill_viridis_c(option = vars_color[x], alpha = 0.9, na.value = "transparent"
          , limits = c(vars_llim[x], vars_ulim[x])
      ) +
      geom_sf(data = units, alpha = 0, lwd = 1, color = "black") +
      labs(
          title = vars_title[x]
          , fill = vars_title[x]
        ) +
      xlab("") +
      ylab("") +
      scale_x_continuous(expand = c(0, 0)) +
      scale_y_continuous(expand = c(0, 0)) +
      theme_bw() +
      theme(
        legend.position = "bottom"
        , legend.direction = "horizontal"
        , legend.margin = margin(0,0,0,0)
        , legend.text = element_text(size = 7)
        , legend.title = element_text(size = 7)
        , axis.text.y = element_text(size = 7)
        , axis.text.x = element_text(size = 7, angle = 35, hjust = 0.7)
        , panel.border = element_blank()
      ) 
  )
}
```

## Spatial groups and HLI

See [this section](#hli) for discussion on HLI. HLI is a proxy for aspect-driven solar heating.

Higher values (i.e. closer to 1) represent warmer conditions while lower values (i.e. closer to 0) represent cooler conditions.

```{r}
# load basemap
hli_basemap <- plt_fn(which(vars=="hli"))
# openings
hli_basemap +
  geom_sf(
    data = all_groups_topo %>% dplyr::filter(group == "Openings")
    , fill = "white"
    , color = "firebrick"
    , lwd = 0.5
  ) +
  geom_sf(
    data = all_groups_topo %>% dplyr::filter(group == "Reserves")
    , fill = "black"
    , color = "black"
  ) +
  labs(
    subtitle = "Reserves = black, Openings = white"
  )
```

## Summary of groups and HLI

```{r, results='asis'}
#summary
all_groups_topo %>% sf::st_set_geometry(NULL) %>% 
  dplyr::group_by(group) %>%
  dplyr::summarise(
    N = n()
    , mean_hli = mean(hli, na.rm = T) 
    , min_hli = min(hli, na.rm = T) 
    , max_hli = max(hli, na.rm = T) 
    , hli_25 = quantile(hli, 0.25)
    , hli_75 = quantile(hli, 0.75)
    
  ) %>%  
  kableExtra::kable(
    caption = "Summary statistics by group type"
    , col.names = c(
      "Group Type"
      , "# Groups"
      , "Mean HLI"
      , "Min. HLI"
      , "Max. HLI"
      , "25% HLI"
      , "75% HLI"
      )
    , digits = 2
  ) %>% 
  kableExtra::kable_styling(font_size = 12) %>% 
  kableExtra::column_spec(1, bold = TRUE, width = "18em")
```

## Quartiles of HLI

```{r}
all_groups_topo_fnl <- all_groups_topo %>% 
  dplyr::group_by(group) %>% 
  dplyr::mutate(
    hli_group_percent_rank = dplyr::percent_rank(hli)
    , hli_group_qrtl = dplyr::case_when(
      hli_group_percent_rank <= .25 ~ "Coolest"
      , hli_group_percent_rank <= .5 ~ "Med. Cool"
      , hli_group_percent_rank <= .75 ~ "Med. Warm"
      , hli_group_percent_rank <= 1 ~ "Warmest"
      , TRUE ~ "ERROR"
    )  
  ) %>% 
  dplyr::mutate(
    hli_group_qrtl = factor(
        hli_group_qrtl
        , ordered = TRUE
        , levels = c("Coolest", "Med. Cool", "Med. Warm", "Warmest")
      )
  ) %>% 
  dplyr::ungroup() %>% 
  dplyr::mutate(
    hli_overall_percent_rank = dplyr::percent_rank(hli)
    , hli_overall_qrtl = dplyr::case_when(
      hli_overall_percent_rank <= .25 ~ "Coolest"
      , hli_overall_percent_rank <= .5 ~ "Med. Cool"
      , hli_overall_percent_rank <= .75 ~ "Med. Warm"
      , hli_overall_percent_rank <= 1 ~ "Warmest"
      , TRUE ~ "ERROR"
    )  
  ) %>% 
  dplyr::mutate(
    hli_overall_qrtl = factor(
        hli_overall_qrtl
        , ordered = TRUE
        , levels = c("Coolest", "Med. Cool", "Med. Warm", "Warmest")
      )
  ) %>% 
  sf::st_make_valid()

sf::st_write(all_groups_topo_fnl, "../data/all_groups_topo_fnl.gpkg", append = FALSE)
```

### HLI Quartiles by treatment type

```{r}
all_groups_topo_fnl %>% 
  sf::st_set_geometry(NULL) %>% 
  dplyr::group_by(group, hli_group_qrtl) %>% 
  dplyr::summarise(
    mean_hli = mean(hli, na.rm = T)     
  ) %>% 
ggplot(.) + 
  geom_col(mapping = aes(y = mean_hli, x = hli_group_qrtl, fill = group), width = 0.7) +
  geom_text(
    aes(x = hli_group_qrtl, y = mean_hli, label = scales::comma(mean_hli, accuracy = .01))
    , color = "black", size = 4
    , position = position_dodge(0.9)
    , vjust = -0.1
  ) +
  facet_grid(.~group) +
  labs(
      title = "Mean HLI of Groups by treatement for cool-warm sites"
    ) +
  xlab("Cool-Warm Site") +
  ylab("Mean HLI") +
  scale_fill_manual(values = c("gray80", "black")) +
  theme_bw() +
  theme(
    legend.position = "none"
  )
```


There is a difference in the mean HLI for the cool and medium cool groups across the opening and reserve treatments.

### HLI Quartiles overall

```{r}
temp_dta <- all_groups_topo_fnl %>% 
  sf::st_set_geometry(NULL) %>% 
  dplyr::group_by(hli_overall_qrtl) %>% 
  dplyr::summarise(
    mean_hli = mean(hli, na.rm = T)     
    , n = n()
  )  %>% 
  dplyr::mutate(nlab = paste0("n=",n))
ggplot(temp_dta) + 
  geom_col(mapping = aes(y = mean_hli, x = hli_overall_qrtl), fill = "navy", width = 0.7) +
  geom_text(
    aes(x = hli_overall_qrtl, y = mean_hli, label = scales::comma(mean_hli, accuracy = .01))
    , color = "black", size = 4
    , position = position_dodge(0.9)
    , vjust = -0.7
  ) +
  geom_text(
    aes(x = hli_overall_qrtl, y = mean_hli, label = nlab)
    , color = "black", size = 2.5
    , position = position_dodge(0.9)
    , vjust = -0.1
  ) +
  labs(
      title = "Mean HLI of Groups by treatement for cool-warm sites"
    ) +
  xlab("Cool-Warm Site") +
  ylab("Mean HLI") +
  scale_fill_manual(values = c("gray80", "black")) +
  theme_bw() +
  theme(
    legend.position = "none"
  )
```

### Counts by cool-warm and treatment group

```{r}
all_groups_topo_fnl %>% 
  sf::st_set_geometry(NULL) %>% 
  dplyr::group_by(group, hli_overall_qrtl) %>% 
  dplyr::summarise(n = n()) %>% 
ggplot(.) +
  geom_bar(aes(x = hli_overall_qrtl, y = n, fill = group, label = n)
    , stat = "identity", width = 0.7) +
  geom_text(aes(x = hli_overall_qrtl, y = n, fill = group, label = n)
    , size = 4, position = position_stack(vjust = 0.5), color = "white") +
  geom_text(data = temp_dta
    , aes(x = hli_overall_qrtl, y = n, label = n)
    , color = "black", size = 5
    , position = position_dodge(0.9)
    , vjust = -0.2
  ) +
  labs(
      title = "Count of treatment groups for cool-warm sites"
    ) +
  xlab("Cool-Warm Site") +
  ylab("Count Groups") +
  scale_fill_manual(values = c("gray80", "black")) +
  theme_bw() +
  theme(
    legend.position = "top"
    , legend.title = element_blank()
  )
```


```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}
remove(list = ls()[grep("temp_",ls())])
gc()
```

## Map Cool and Warm Sites

```{r, results=FALSE}
p_hli_fn <- function(x){
  ggplot() + 
    geom_sf(
      data = units
      , lwd = 0.8
      , alpha = 0
      , color = "black" 
    ) +
    geom_sf(
      data = all_groups_topo_fnl %>% dplyr::filter(group == x)
      , aes(fill = hli_overall_qrtl)
      , lwd = 0
    ) +
    scale_fill_viridis_d(option = "turbo", alpha = 0.9) +
    labs(
      title = paste(x, "- Heat Load Index")
      , subtitle = "harvest treatment units in black"
    ) +
    theme_bw() + 
    theme(
      legend.position = "top"
      , legend.title = element_blank()
      , axis.text.y = element_text(size = 7)
      , axis.text.x = element_text(size = 7, angle = 35, hjust = 0.7)
      , panel.border = element_blank()
    ) 
}
c(unique(all_groups_topo_fnl$group)) %>% 
  purrr::map(p_hli_fn)
```

## Characteristics of Cool-Warm Sites

### Summary statistics

```{r}
# vars to aggregate
vars_temp <- c(
  "hli"
  , "area_m2"
)
vars_lab_temp <- c(
  "HLI"
  , latex2exp::TeX("Area ($m^{2}$)")
)
# aggregate data to daily level
# named list of functions
summary_fns <- list(
  mean = ~mean(.x, na.rm = TRUE)
  # , median = ~median(.x, na.rm = TRUE)
  , min = ~min(.x, na.rm = TRUE)
  , max = ~max(.x, na.rm = TRUE)
  # , N = ~sum(ifelse(is.na(.x), 0, 1))
)

all_groups_topo_fnl %>%
  sf::st_set_geometry(NULL) %>%
  dplyr::group_by(hli_overall_qrtl) %>%
  dplyr::summarise(
    dplyr::across(
      tidyselect::all_of(vars_temp)
      , summary_fns
      , .names = "{.col}_{.fn}"
    )
  ) %>%
  dplyr::ungroup() %>%
  kableExtra::kable(
    caption = "Summary statistics by group type"
    , col.names = c(
      ""
      , "Mean HLI"
      , "Min. HLI"
      , "Max. HLI"
      , "Mean Area m sq."
      , "Min. Area m sq."
      , "Max. Area m sq."
      
    )
    , digits = 2
  ) %>% 
  kableExtra::kable_styling(font_size = 12) %>% 
  kableExtra::column_spec(1, bold = TRUE, width = "18em")
```

### HLI Distribution

```{r, results=FALSE}
p_fn <- function(my_var) {
  (
    ggplot() + 
      geom_density(data = all_groups_topo_fnl,
        mapping = aes_string(x = my_var, fill = "hli_overall_qrtl")
        , color = NA
      ) + 
      facet_grid(hli_overall_qrtl~.) +
      scale_fill_viridis_d(option = "turbo", alpha = 0.9) +
      scale_x_continuous(breaks = scales::extended_breaks(n=10)) +
      xlab(vars_lab_temp[which(vars_temp == my_var)]) +
      ylab("Density") +
      labs(
        title = vars_lab_temp[which(vars_temp == my_var)]
        , subtitle = "distribution by cool-warm site type"
      ) +
      theme_bw() + 
      theme(
        legend.position = "top"
        , legend.title = element_blank()
        , axis.text.y = element_text(size = 7)
        , axis.text.x = element_text(size = 7, angle = 35, hjust = 0.7)
        , panel.border = element_blank()
      )
  )
}
# call function
vars_temp %>% 
  purrr::map(p_fn) 
```

### Length:Width ratio of group selection shape

```{r, eval=FALSE, include=FALSE}
# get bbox output as data frame
for (i in 1:nrow(all_groups_topo_fnl)){
  dta_temp <- data.frame(
    shape_id = all_groups_topo_fnl$shape_id[i]
    , group = all_groups_topo_fnl$group[i]
    , xmin = sf::st_bbox(all_groups_topo_fnl$geom[i])[1]
    , xmax = sf::st_bbox(all_groups_topo_fnl$geom[i])[3]
    , ymin = sf::st_bbox(all_groups_topo_fnl$geom[i])[2]
    , ymax = sf::st_bbox(all_groups_topo_fnl$geom[i])[4]
  )  
  if(i==1){
    dta_join_temp <- dta_temp
  }else{
    dta_join_temp <- rbind(dta_join_temp, dta_temp)
  }
  remove(dta_temp)
}

# join to main data
all_groups_topo_fnl <- all_groups_topo_fnl %>%
  dplyr::left_join(dta_join_temp, by=c("shape_id"="shape_id", "group"="group")) %>% 
  dplyr::mutate(
    xlength_m = xmax - xmin
    , ylength_m = ymax - ymin
    , length_width_ratio = ylength_m/xlength_m
    , ns_ew_ratio = ylength_m/xlength_m
  )
```

